A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-03-15 survey.tex
a few corrections / arbitrator precision

On the other hand, bots such as Nova or BroodwarBotQ (BBQ) only use a hierarchical organization for {\em micro-management} (controlling the attack units), but use a decentralized organization for the rest of the bot. In Nova and BBQ, there is a collection of modules that control different aspects of the game (workers, production, construction, etc.). These modules can all send actions directly to StarCraft. In Nova those modules coordinate mostly through writing data in a shared blackboard, and in BBQ they coordinate only when they have to use a shared resource (unit) by means of an arbitrator.

@@

\begin{figure}[t]
	%\vspace*{-1ex}
    \centering
    \includegraphics[width=0.95\columnwidth]{figures/cig2012-ResultsRound90.pdf}
	%\vspace*{-1ex}
    \caption{Evolution of the win percentage of each bot participating in the
CIG 2012 competition}
    \label{fig:cig2012-results}
\end{figure}

@@

One of the major conclusions from the results of the StarCraft competitions, is the lack of adaptation of bots.  Some switch between different build-orders, but do not fully adapt their strategy. No bot is capable of observing the opponent and autonomously synthesize a good plan from scratch to counter the opponent strategy. In this section we analyzed this claim quantitatively. We analyzed the replays from the 2011 and 2012 AIIDE competitions (shown in Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} respectively). Specifically, we analyzed the way bots choose their openings depending on which other bot they are playing against. Given that there is no dominant strategy in StarCraft, and it is necessary to see what the opponent is doing in order to determine the best opening, we would expect bots to change their openings depending on which other bot they are playing (since each bot uses a different strategy, or set of strategies). Using clustering, we identified the most common openings in all the replays from the 2011 and 2012 competitions (each one shown with a different color in the figures). No data is shown for the ItayUnvermind bot, since its opening did not match significantly with any of the ones used in our study.

Specifically, we identified the following openings (Most of the following names are specific building or units of StarCraft, for a better comprehension we refer the reader to \footnote{\url{http://wiki.teamliquid.net/starcraft/Category:Strategies}}):
\begin{itemize}
  \item Protoss openings:
  \begin{itemize}
    \item \emph{two\_gates}: Build two \emph{Gateways} and keep training \emph{Zealots} (basic contact attack ground unit), this is the quickest way to apply pressure.
    \item \emph{fast\_dt}: Produce \emph{Dark Templars} (technologically advanced stealth ground unit) as soon as possible, sacrificing early game power for a technological advance, hard-countered by detectors technology.
    \item \emph{templar}: Train \emph{High Templars} (technologically advanced zone attack unit) as fast as possible, same as above, less deadly but is less easily countered.
    \item \emph{speedzeal}: Train \emph{Zealots} and research attack and speed upgrades as soon as possible, some early game power transitioning into late game tech.
    \item \emph{corsair}: Produce \emph{Corsairs} (air-air flying unit) as soon as possible and then transition into training \emph{Dark Templars} (safe from Zerg's flying detectors thanks to \emph{Corsairs}), \emph{Reavers} (ground artillery unit) or \emph{Dragoons}. Weak early game.
    \item \emph{nony}: Build three \emph{Gateways} and massive training of \emph{Dragoons}. Slower than \emph{two\_gates} but still some early game (ranged) power.
    \item \emph{reaver\_drop}: Train \emph{Reavers} as soon as possible to be able to do drops (air transport of artillery units).
  \end{itemize}
  \item Terran openings:
  \begin{itemize}
    \item \emph{bio}: Produce a large army of \emph{Marines} (basic ranged ground unit) and \emph{Medics} (can heal biological units). Quickest way to apply pressure.
    \item \emph{rax\_fe}: Take the closest ``natural expansion'' as soon as possible. This provides a big economic boost in the mid game by sacrificing some early game power.
    \item \emph{two\_facto}: Build two \emph{Factories} and keep producing \emph{Tanks} (ground artillery unit). Vulnerable while building up to it and then very powerful on ground.
    \item \emph{vultures}: Produce mainly \emph{Vultures} (fast ground ranged unit, excels against small units) and research mines. Quicker to reach (technologically) and build than tanks, can transition into tanks.
    \item \emph{air}: Produce \emph{Wraiths} (ranged flying units) as soon as possible for an air attack. Vulnerable to anti-air openings or quick rushes.
    \item \emph{drop}: Train \emph{Dropships} (flying transports) as soon as possible to be able to do (mostly tanks or marines) drops, leveraging efficient tactics.
  \end{itemize}
  \item Zerg openings:
  \begin{itemize}
    \item \emph{speedlings}: Train \emph{Zerlings} (basic cheap, fast, ground contact attack unit) and research speed upgrade as soon as possible. Quickest way to apply pressure.
    \item \emph{fast\_mutas}: Produce mainly \emph{Mutalisks} (ranged flying units). Vulnerable in the early game while gathering gas and researching the technology.
    \item \emph{mutas}: Expand two times for a stronger economy before massive training of \emph{Mutalisks}. Slower but more powerful build-up than above.
    \item \emph{lurkers}: Train \emph{Lurkers} (ground, stealth artillery unit) as soon as possible to benefit from their (advanced technology) zone attack and cloak ability. 
    \item \emph{hydras}: Massive production of \emph{Hydralisks} (ground ranged unit). Much quicker to reach technologically than \emph{Lurkers} and can transition into them.
  \end{itemize}
\end{itemize}

>>>>

On the other hand, bots such as Nova or BroodwarBotQ (BBQ) only use a hierarchical organization for {\em micro-management} (controlling the attack units), but use a decentralized organization for the rest of the bot. In Nova and BBQ, there is a collection of modules that control different aspects of the game (workers, production, construction, etc.). These modules can all send actions directly to StarCraft. In Nova those modules coordinate mostly through writing data in a shared blackboard, and in BBQ they coordinate only when they have to use a shared resource (unit) by means of an arbitrator: a bidding market and broker for settling units control, military and civilian groups/task forces bid for units proportionally to their usefulness and the task importance.

@@

\begin{figure}[t]
	%\vspace*{-1ex}
    \centering
    \includegraphics[width=\columnwidth]{figures/cig2012-ResultsRound90.pdf}
	%\vspace*{-1ex}
    \caption{Evolution of the win percentage of each bot participating in the
CIG 2012 competition}
    \label{fig:cig2012-results}
\end{figure}

@@

One of the major conclusions from the results of the StarCraft competitions, is the lack of adaptation of bots.  Some switch between different build-orders, but do not fully adapt their strategy. No bot is capable of observing the opponent and autonomously synthesize a good plan from scratch to counter the opponent strategy. In this section we analyzed this claim quantitatively using the tools of \cite{SynnaeveOpeningCig11}. We analyzed the replays from the 2011 and 2012 AIIDE competitions (shown in Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} respectively). We analyzed the way bots choose their openings depending on which other bot they are playing against. Given that there is no dominant strategy in StarCraft, and it is necessary to see what the opponent is doing in order to determine the best opening, we would expect bots to change their openings depending on which other bot they are playing (since each bot uses a different strategy, or set of strategies). Using clustering, we identified the most common openings in all the replays from the 2011 and 2012 competitions (each one shown with a different color in the figures). No data is shown for the ItayUnvermind bot, since its opening did not match significantly with any of the ones used in our study (extracted from humans pro-gamers).

Specifically, we identified the following openings (for a better comprehension of strategies, buildings or units of StarCraft, we refer the reader to Teamliquid's wiki\footnote{\url{http://wiki.teamliquid.net/starcraft/Category:Strategies}}):
\begin{itemize}
  \item Protoss openings:
  \begin{itemize}
    \item \emph{two\_gates}: Build two \emph{Gateways} and keep training \emph{Zealots} (basic contact attack ground unit), this is the quickest way to apply pressure.
    \item \emph{fast\_dt}: Produce \emph{Dark Templars} (technologically advanced stealth ground unit) as soon as possible, sacrificing early game power for a technological advance, hard-countered by detectors technology.
    \item \emph{templar}: Train \emph{High Templars} (technologically advanced zone attack unit) as fast as possible, same as above, less deadly but is less easily countered.
    \item \emph{speedzeal}: Train \emph{Zealots} and research attack and speed upgrades as soon as possible, some early game power transitioning into late game tech.
    \item \emph{corsair}: Produce \emph{Corsairs} (air-air flying unit) as soon as possible and then transition into training \emph{Dark Templars} (safe from Zerg's flying detectors thanks to \emph{Corsairs}), \emph{Reavers} (ground artillery unit) or \emph{Dragoons}. Weak early game.
    \item \emph{nony}: Build three \emph{Gateways} and massive training of \emph{Dragoons}. Slower than \emph{two\_gates} but still some early game (ranged) power.
    \item \emph{reaver\_drop}: Train \emph{Reavers} as soon as possible to be able to do drops (air transport of artillery units).
  \end{itemize}
  \item Terran openings:
  \begin{itemize}
    \item \emph{bio}: Produce a large army of \emph{Marines} (basic ranged ground unit) and \emph{Medics} (can heal biological units). Quickest way to apply pressure.
    \item \emph{rax\_fe}: Take the closest ``natural expansion'' as soon as possible. This provides a big economic boost in the mid game by sacrificing some early game power.
    \item \emph{two\_facto}: Build two \emph{Factories} and keep producing \emph{Tanks} (ground artillery unit). Vulnerable while building up to it and then very powerful on ground.
    \item \emph{vultures}: Produce mainly \emph{Vultures} (fast ground ranged unit, excels against small units) and research mines. Quicker to reach (technologically) and build than tanks, can transition into tanks.
    \item \emph{air}: Produce \emph{Wraiths} (ranged flying units) as soon as possible for an air attack. Vulnerable to anti-air openings or quick rushes.
    \item \emph{drop}: Train \emph{Dropships} (flying transports) as soon as possible to be able to do (mostly tanks or marines) drops, leveraging efficient tactics.
  \end{itemize}
  \item Zerg openings:
  \begin{itemize}
    \item \emph{speedlings}: Train \emph{Zerlings} (basic cheap, fast, ground contact attack unit) and research speed upgrade as soon as possible. Quickest way to apply pressure.
    \item \emph{fast\_mutas}: Produce mainly \emph{Mutalisks} (ranged flying units). Vulnerable in the early game while gathering gas and researching the technology.
    \item \emph{mutas}: Expand two times for a stronger economy before massive training of \emph{Mutalisks}. Slower but more powerful build-up than above.
    \item \emph{lurkers}: Train \emph{Lurkers} (ground, stealth artillery unit) as soon as possible to benefit from their (advanced technology) zone attack and cloak ability. 
    \item \emph{hydras}: Massive production of \emph{Hydralisks} (ground ranged unit). Much quicker to reach technologically than \emph{Lurkers} and can transition into them.
  \end{itemize}
\end{itemize}
